The pragmatics of perspective-taking

### 2018-05-13 writing/journal_manuscript/manuscript.tex
Finish up exp. 1; put all figures in

 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=1]{figures/resultsFig_v1.pdf}
 \vspace{-.5cm}
 \caption{Listener errors with scripted and unscripted instructions. Error bars are bootstrapped 95\% confidence intervals.}
 \vspace{-.5cm}
 \label{fig:exp1results}
 \end{center}
 \end{figure}

@@

Our scripted condition successfully replicated the results of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} with even stronger effects: listeners attempted to move the hidden object on approximately 50\% of critical trials. However, listeners attempted to move the hidden object less than half as often when their partner was able to produce natural referring expressions, $p_1 = 0.51, p_2 = 0.20, \chi^2(1) = 48.5, p < 0.001$ (Fig. \ref{fig:exp1results}A). We also found substantial heterogeneity in errors rates across items, which diverge significantly from a uniform distribution in both scripted ($\chi^2=55, p < 0.001$) and unscripted ($\chi^2=36, p <0.001$) conditions under a non-parametric $\chi^2$ test (Fig. \ref{fig:exp1results}B). Still, listeners in the unscripted condition made fewer errors for nearly every item, some more dramatically than others. A maximal mixed-effects logistic regression with random intercepts for each dyad and random slopes and intercepts for each item shows a significant effect of condition ($z = 2.856, p = 0.004$). %For items where participants 
%For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.''

@@

Even if participants in the unscripted condition make fewer actual errors, they may still be \emph{considering} the hidden object just as often on trials where they go on to make correct responses. As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis. Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult and unreliable to measure. 
So we used as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
\footnote{We preregistered as our dependent variable the mean \emph{amount of time} spent hovering over hidden distractor, but because this DV was exactly zero for the majority of trials and a large, highly variable number for the rest of trials, we decided post-hoc that the mean was not as meaningful as the binarized version. Still, regression results are unchanged when conducted on log-transformed hover times.} 
We found that listeners in the \emph{scripted} condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. In the unscripted condition, however, there was no difference from the baseline. We found a significant interaction in a maximal mixed-effect model with dyad-level and object-level random intercepts and a dyad-level random effect for the difference from baseline ($z = -2.8, p = 0.004$; Fig. \ref{fig:exp1results}C). %predicted that even when participants went on to make a correct decision, they would be 
%We define the decision window as the span of time between  their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. 
% Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor. %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. 
%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

@@

Next, we test whether these improvements in listener performance are in fact due to more informative speaker behavior in the unscripted condition. %By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
%First, on filler trials in both conditions, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. For example, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
%Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
%If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
%That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
%We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
The simplest measure of speaker informativity is the raw number of   words used in referring expressions. Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5, p < 0.001$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.

\todo[inline]{Update following 2 paragraphs with details once finalized} 
For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances. We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects. Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ). In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (Fig. \ref{fig:exp1results}D, $b = AAA, t = BBB, p < CCC$). The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this scenario.

To what extent does differential informativity of utterances account for the heterogeneity in error? %In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit. We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$. This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances. 
We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, (\emph{coefficient STAT STAT STAT}) and more likely to make errors when the distractor fit is higher, (\emph{coefficient STAT STAT STAT}). Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = XXX, p < YYY$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

@@

Our primary is the length (in words) of their naturally produced referring expressions in different contexts.

As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color).

>>>>

 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=1]{figures/Exp1_listener_results.pdf}
 \vspace{-.5cm}
 \caption{Listener results for Exp.~1. (A) Errors distributions with scripted and unscripted instructions -- participants in the unscripted condition made significantly fewer errors. (B) Even when made the correct response, listeners in the scripted condition were more likely to hover over the distractor relative to baseline while the unscripted condition shows no difference. Error bars are bootstrapped 95\% confidence intervals.}
 \vspace{-.5cm}
 \label{fig:exp1listener}
 \end{center}
 \end{figure}

@@

Our scripted condition successfully replicated the results of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} with even stronger effects: listeners attempted to move the hidden object on approximately 50\% of critical trials. However, listeners attempted to move the hidden object less than half as often when their partner was able to produce natural referring expressions, $p_1 = 0.51, p_2 = 0.20, \chi^2(1) = 48.5, p < 0.001$ (Fig. \ref{fig:exp1listener}A). We also found substantial heterogeneity in errors rates across items, which diverge significantly from a uniform distribution in both scripted ($\chi^2=55, p < 0.001$) and unscripted ($\chi^2=36, p <0.001$) conditions under a non-parametric $\chi^2$ test (see Supplemental Fig. \ref{fig:error_heterogeneity}). Still, listeners in the unscripted condition made fewer errors for nearly every item, some more dramatically than others. A maximal mixed-effects logistic regression with random intercepts for each dyad and random slopes and intercepts for each item shows a significant effect of condition ($z = 2.856, p = 0.004$). %For items where participants 
%For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.''

@@

Even if participants in the unscripted condition make fewer actual errors, they may still be \emph{considering} the hidden object just as often on trials where they go on to make correct responses. As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis. Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult and unreliable to measure. 
So we used as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
\footnote{We preregistered as our dependent variable the mean \emph{amount of time} spent hovering over hidden distractor, but because this DV was exactly zero for the majority of trials and a large, highly variable number for the rest of trials, we decided post-hoc that the mean was not as meaningful as the binarized version. Still, regression results are unchanged when conducted on log-transformed hover times.} 
We found that listeners in the \emph{scripted} condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. In the unscripted condition, however, there was no difference from the baseline. We found a significant interaction in a maximal mixed-effect model with dyad-level and object-level random intercepts and a dyad-level random effect for the difference from baseline ($z = -2.8, p = 0.004$; Fig. \ref{fig:exp1listener}B). %predicted that even when participants went on to make a correct decision, they would be 
%We define the decision window as the span of time between  their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. 
% Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor. %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. 
%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

@@

 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=1]{figures/Exp1_speaker_results.pdf}
 \vspace{-.5cm}
 \caption{Speaker results for Exp.~1. (A) While scripted utterances were judged to fit target and distractor roughly equally, speakers in the unscripted condition naturally produced utterances that fit the target much better than the distractor. (B) The extent to which an utterance fits the target more than the distractor is highly predictive of error rates at an item-by-item level. All error bars are bootstrapped 95\% confidence intervals.}
 \vspace{-.5cm}
 \label{fig:exp1speaker}
 \end{center}
 \end{figure}

Next, we test whether these improvements in listener performance are in fact due to more informative speaker behavior in the unscripted condition. %By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
%First, on filler trials in both conditions, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. For example, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
%Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
%If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
%That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
%We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
The simplest measure of speaker informativity is the raw number of   words used in referring expressions. Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.
\todo[inline]{We can remove the above analysis for space if necessary}

For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 123 unique labels after including scripted utterances. We recruited 20 judges on Amazon Mechanical Turk to rate whether each label was true of the target and hidden distractor objects. They were shown objects in the context of the full grid (with no occlusions) such that they could feasibly judge spatial or relative references like ``bottom block.'' Responses were given on a slider with endpoints labeled ``strongly agree'' and ``strongly disagree.''  After excluding 4 judges for guessing with response times $< 1s$, inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.54 (95\% CI = [0.47, 0.61])$. Because all judges rated all utterances for both target and distractor, we could compute the \emph{informativity} of an utterance as the difference score between target and distractor responses.

Our primary measure of interest is the difference in informativity across conditions. Due to the complexity of the multiple partially crossed levels of variability in our data, standard mixed-effects models were not appropriate. Instead, we used a multi-stage, non-parametric bootstrap to compute 95\% confidence intervals. We found that speakers in the unscripted condition systematically produced utterances with higher informativity ratings than the scripted utterances ($m = 0.5$, 95\% bootstrapped CI = $[0.23, 0.78], p < .001$; see Fig. \ref{fig:exp1speaker}A). Scripted labels fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this context.

Finally, we consider the link between our speaker and listener results. If listeners have a well-tuned pragmatic expectation about the speaker's likely level of informativity, they may rationally neglect the speaker's perspective and simply pick the object that fits the description best. To what extent does the informativity of speaker utterances account for heterogeneity in listener errors across items in the two conditions? As predicted by a perspective-negligence account, we found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a good fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig. \ref{fig:exp1speaker}B). %  We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when informativity is higher. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = XXX, p < YYY$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
\todo[inline]{rxdh: I think this bootstrapped correlation stat at the objectSet-condition level makes this point well enough for our purposes (it's a pretty huge effect...), but I could also do a fancier thing boostrapping logistic coefficients at the trial-by-trial level if necessary. }
\todo[inline]{rxdh: also, is it worth having a short appendix with more details on the multi-stage bootstrapping procedure? it's very straightforward but takes a few sentences to explain\dots}

@@

\subsection{Results}
 \begin{figure}[t!]
 \begin{center}
 \vspace{-.25cm}
 \includegraphics[scale=1.25]{figures/Exp2_results.pdf}
 \vspace{-.5cm}
 \caption{Results for Exp.~2. (A) Speakers used significantly more words when occlusions were present, and effect is even greater than the standard pragmatic effect of being in a close vs. far context (B) Utterances broken out by feature mentioned. All error bars are bootstrapped 95\% confidence intervals.}
 \vspace{-.5cm}
 \label{fig:exp2results}
 \end{center}
 \end{figure}

Our primary is the length (in words) of their naturally produced referring expressions in different contexts. \ref{fig:exp2results}A

As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color). \ref{fig:exp2results}B
